---
title: "Cap Dap with UV Data"
author: "Anna"
date: "11/29/2022"
output: html_document
---

In this file I will be working with data that includes D, NL, and UV treatments (darkness, natural light, and UV light, respectively). 

First, set things up:
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

The data have been updated to include new dates and UV data:
```{r}
DF <- read.csv(here("Data", "cap_color_data_final.csv"))
```

I want to observe my data to make sure all variables are included and there isn't anything funky going on
```{r}
str(DF)
```
I see that some of the variables are listed as character data, so I want to convert those into factors.
```{r}
DF$treatment <- as.factor(DF$treatment)
```

Before plotting any data, I will look at a summary of the data to see the spread and see if anything sticks out that may be wrong with it.
```{r}
summary(DF)
```
In the "a" column, the max is 44 which seems odd. From working with this data for the past few weeks, I know that 44 is an extremely high number on this scale so it is probably a typo. I feel confident that value was supposed to be 4 instead of 44 so I will make that change.
```{r}
DF[267, 3] = 4
```

For right now, I just want to look at the copepod samples from the November 3rd group, so I will create a new data frame with only this data after the copepods have been in their light treatments for 7 days. 
```{r}
Nov3 <- DF[c(468:706), c(2:5)]
```

Let's plot this data to get a general idea of the spread of data. First, let's look at **Luminosity** 
```{r}
ggplot(Nov3, aes(x = before_after, y = L)) +
geom_boxplot() +
xlab("Treatment") +
ylab("Luminosity") +
theme_bw() +
facet_wrap(~treatment) +
scale_x_discrete(limits = c("before", "after"))
```



