---
title: "NL vs. D vs. UV"
author: "Anna"
date: "11/30/2022"
output: html_document
---

First, set things up:
```{r}
rm(list=ls())
library(here)
library(tidyverse)
library(ggfortify)
```

The data have been updated to include new dates and UV data:
```{r}
DF <- read.csv(here("Data", "cap_color_data_final.csv"))
```

I want to observe my data to make sure all variables are included and there isn't anything funky going on
```{r}
str(DF)
```

Now I want to make a new data frame to grab just the variables I am interested in 
```{r}
Nov3 <- DF %>%
  filter(date_collected == "11/3/22") %>%
  select(L, a, b, treatment)
```

The "before" group is listed as "NA" for treatment, so I want to change that 
```{r}
Nov3$treatment <- Nov3$treatment %>%
  replace_na("control")
```

```{r}
Nov3$treatment <- as.factor(Nov3$treatment)
```


Let's look at the structure of this new data frame now
```{r}
str(Nov3)
```
Let's do a quick plot to look at the general spread of data
```{r}
ggplot(Nov3, aes(x = treatment, y = L)) +
geom_boxplot() +
xlab("Treatment") +
ylab("Luminosity") +
theme_bw() 
```

```{r}
model_L <- glm(L ~ treatment, data = Nov3, family = poisson)
autoplot(model_L, smooth.color = NA)
```

```{r}
anova(model_L, test = "Chisq")
```

Somewhere in these treatments, there is significant difference in luminosity between treatments

```{r}
mean_df <- Nov3 |> 
  group_by(treatment) |>
  summarise(mean_lum = mean(L))
```

```{r}
mean_control <- Nov3 |>
  filter(treatment == "control") |>
  summarise(mean = mean(L))
```


```{r}
ggplot(Nov3, aes(x = treatment, y = L, color = treatment)) +
  geom_point(size = 3, alpha = 0.3) +
  geom_point(data = mean_df, aes(x = treatment, y = mean_lum), shape = "diamond", size = 7, show.legend = FALSE) +
  geom_hline(data = mean_control, aes(yintercept = mean), linetype = 2) +
  theme_bw() +
  xlab("Treatment") +
  ylab("Luminosity (L)") +
  coord_flip()
```

Perform a Tukey test to see where the significant difference lies
```{r}
modelL_tukey <- aov(L ~ treatment, data = Nov3)
TukeyHSD(modelL_tukey, conf.level=.95) 
```
The copepods in the NL and D treatments became significantly more luminous. The UV copepods stayed darker and did not fade significantly from the original color.

Now, to look at the b-value
```{r}
ggplot(Nov3, aes(x = treatment, y = b)) +
geom_boxplot() +
xlab("Treatment") +
ylab("b-value") +
theme_bw() 
```
There does not seem to be any significant difference based on this plot, but let's run the stats
```{r}
model_b_tukey <- aov(b ~ treatment, data = Nov3)
TukeyHSD(model_b_tukey, conf.level=.95) 
```
b-values never reach into the negative values, meaning the copepods are not gaining any blue pigment in any treatments. 

Now to look at the a-value:
```{r}
ggplot(Nov3, aes(x = treatment, y = a)) +
geom_boxplot() +
xlab("Treatment") +
ylab("a-value") +
theme_bw() 
```
And a Tukey test...
```{r}
model_a_tukey <- aov(a ~ treatment, data = Nov3)
TukeyHSD(model_a_tukey, conf.level=.95) 
```

